Dataset Summary

### 2018-08-03 Cleaning & subset.Rmd
Meta analysis of body size at ~100 days

# Export

Body size only

@@

keep age closet t0 10 
```{r}

@@

i <- match(unique(x$biological_sample_id), x$biological_sample_id)

data2 <- x[i,] 
```

Calculate number of reocrds per population
```{r}
data2 %>% 
  group_by(population) %>%
  summarise(
    n_records = n_distinct(biological_sample_id)
  ) -> population_size
hist(population_size$n_records)

sum(population_size$n_records < 10)
```

data %>% 
  filter(
    age_in_days >= 100, 
    parameter_name == "body weight") %>%
  select(biological_sample_id, external_sample_id) %>%
  arrange(external_sample_id) %>%
  data.frame()

data %>% 
  filter(
    age_in_days >= 100, 
    parameter_name == "body weight") %>%
  mutate(age_diff=abs(100-age_in_days)) %>%
  group_by(biological_sample_id) %>%
  filter(age_diff == min(age_diff)) -> x

i <- match(unique(x$biological_sample_id), x$biological_sample_id)

x[i,] -> x

# check only one record per individual

table(x$biological_sample_id) -> z 
z[z!=1]

@@

>>>>

Export for body size only

@@

keep age closet t=100

@@

# still some individuals with multiple records (bc appear under different procedures, so filter to one record)
i <- match(unique(x$biological_sample_id), x$biological_sample_id)
data2 <- x[i,] 
```

Calculate number of reocrds per population

```{r}
data2 %>% 
  group_by(population) %>%
  summarise(
    n_records = n_distinct(biological_sample_id)
  ) -> population_size
hist(population_size$n_records)

sum(population_size$n_records < 5)
```

Now calculate statsics by population:

```{r}
population_stats <- data2 %>% 
  filter(!is.na(data_point)) %>% 
  group_by(population, sex, strain_name, production_center) %>% 
  summarise(
    trait = parameter_name[1],
    x_bar = mean(data_point),
    x_sd = sd(data_point),
    n_ind = n()
  ) %>% 
  ungroup() %>%
  filter(n_ind > 5) %>% 
  # Check both sexes present & filter those missing
  group_by(strain_name, production_center) %>% 
  mutate(
    n_sex = n_distinct(sex)
  ) %>% ungroup() %>%
  filter(n_sex ==2) %>% 
  arrange(strain_name, production_center, sex)
```

Now analyse by population data:

Sanity checks

```{r}
i <- seq(1, nrow(population_stats), by=2)
all(population_stats$sex[i] == "female")
all(population_stats$sex[i+1] == "male")
all(population_stats$strain_name[i+1] == population_stats$strain_name[i] )
all(population_stats$production_center[i+1] == population_stats$production_center[i] )

input <- data.frame(
       n1i = population_stats$n_ind[i], 
       n2i = population_stats$n_ind[i+1], 
       x1i = population_stats$x_bar[i], 
       x2i = population_stats$x_bar[i+1],
       sd1i = population_stats$x_sd[i], 
       sd2i = population_stats$x_sd[i+1])

all(input > 0)
```
Check plot of mean vs variance

```{r}
plot(population_stats$x_bar, population_stats$x_sd, pch=16, log="xy", col = as.factor(population_stats$sex))
```

```{r}
Calc.lnCVR <- function(CMean, CSD, CN, EMean, ESD, EN){
	S2 <- log(ESD) - log(EMean) + 1 / (2*(EN - 1)) - (log(CSD) - log(CMean) + 1 / (2*(CN - 1)))
  return(S2)
}

Calc.var.lnCVR <- function(CMean, CSD, CN, EMean, ESD, EN, Equal.E.C.Corr=T) {
	if(Equal.E.C.Corr==T){
		mvcorr <- cor.test(log(c(CMean, EMean)), log(c(CSD, ESD)))$estimate
		S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * mvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * mvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))
	}
	else{
		Cmvcorr <- cor.test(log(CMean), log(CSD))$estimate
		Emvcorr <- cor.test(log(EMean), (ESD))$estimate
		S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * Cmvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * Emvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))		
	}
	return(S2)
}

Calc.lnVR <- function(CSD, CN, ESD, EN){
	log(ESD) - log(CSD) + 1 / (2*(EN - 1)) - 1 / (2*(CN - 1))
}

Calc.var.lnVR <- function( CN,  EN) {
	1 / (2*(EN - 1)) + 1 / (2*(CN - 1))
}

Calc.lnRR <- function(CMean, CSD, CN, EMean, ESD, EN) {
	log(EMean) - log(CMean)
}

Calc.var.lnRR <- function(CMean, CSD, CN, EMean, ESD, EN) {
	CSD^2/(CN * CMean^2) + ESD^2/(EN * EMean^2)
}
```

```{r}
# library(metafor)
#measures: CVR, VR
# escalc(measure="CVR", 
#       n1i = n1i, n2i = n2i, x1i = x1i, x2i = x2i, sd1i = sd1i, sd2i = sd2i, data = input)
# not working

results <- population_stats[i,] %>% 
  select(strain_name, production_center) %>%
  mutate(
    effect_size_CVR = Calc.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
    sample_variance_CVR = Calc.var.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
    effect_size_VR = Calc.lnVR(CSD = input$sd1i, CN = input$n1i, ESD = input$sd2i, EN = input$n2i),
    sample_variance_VR = Calc.var.lnVR(CN = input$n1i, EN = input$n2i),
    effect_size_RR = Calc.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
    sample_variance_RR = Calc.var.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
    err = as.factor(seq_len(n()))
    )

@@

## Compare CVR

```{r}
plot(results$effect_size_CVR, 1/sqrt(results$sample_variance_CVR), ylab="sample precision", xlab="effect size", pch=16, xlim=c(-0.7, 0.7), col=as.factor(results$production_center))
abline(v=0)
```

```{r}
metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR, random = list(~1| strain_name, ~1|production_center, ~1|err), data = results)
```

## Compare VR

```{r}
plot(results$effect_size_VR, 1/sqrt(results$sample_variance_VR), ylab="sample precision", xlab="effect size", pch=16, xlim=c(-0.8, 0.8), col=as.factor(results$production_center))
abline(v=0)
```

```{r}
metafor::rma.mv(yi = effect_size_VR, V = sample_variance_VR, random = list(~1| strain_name, ~1|production_center, ~1|err), data = results)
```

## Compare Means

```{r}
plot(results$effect_size_RR, 1/sqrt(results$sample_variance_RR), ylab="sample precision", xlab="effect size", pch=16, xlim=c(-0.5, 0.5), col=as.factor(results$production_center))
abline(v=0)
```

```{r}
metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR, random = list(~1| strain_name, ~1|production_center, ~1|err), data = results)
```
